---
title: "JCM Revision 1: PD on Facebook"
author: "Jeanne"
date: "2025-05-27"
output: html_document
---

# 3.1 Recruitment and Facebook Participation

## Enrollment flowchart
```{r}
# Load constants for API calls
source("constants.R")
# Load supporting functions
source("functions.R")

# Load participant redcap data
df_attempts = fetch_report(url, participants_token, participants_table1)
```

Calculate number of people (total and subgroup) at different stages of enrollment
```{r}
df_success <- clean_and_transform_table1_data(df_attempts)

# Define conditions for different counts
### Assume anyone we know to be dx-ed with PD 
### or who we don't know to be a caregiver to have PD 
### given we used PD recontact lists
### The rest are explicitly by reporting
pd_condition <- "dx_type == 'Parkinsons disease' | (is.na(dx_type) & (caregiver == 'No' | is.na(caregiver)))"
et_condition <- "dx_type == 'Essential Tremor'"
parkinsonism_condition <- "dx_type == 'Parkinsonism'"
caregiver_condition <- "caregiver == 'Yes' & (pd_yesno == 'Caregivers' | is.na(pd_yesno))"

# Successful contact totals for different subsets
total_attempts <- nrow(df_attempts)
total_contact <- nrow(df_success)

# Total number of people who agreed to participate
df_agree <- subset(df_success, contact_refused == 'No')
total_agree <- nrow(df_agree)

# Total number of people enrolled
df_enrolled <- subset(df_agree, !is.na(cf_dt) | !is.na(enrollment_date))
total_enroll <- nrow(df_enrolled)

# People who withdrew
df_withdrew <- subset(df_enrolled, withdrawn == 'Yes')
total_withdrew <- nrow(df_withdrew)

# People who completed the study (at least the interview)
df_interviewed <- subset(df_enrolled, withdrawn != 'Yes' | is.na(withdrawn))
total_interview <- nrow(df_interviewed)
interview_pd_count <- count_individuals(df_interviewed, pd_condition)
interview_et_count <- count_individuals(df_interviewed, et_condition)
interview_parkinsonism_count <- count_individuals(df_interviewed, parkinsonism_condition)
interview_caregivers_count <- count_individuals(df_interviewed, caregiver_condition)

# People whose Facebook data we acquired
df_facebook <- subset(df_interviewed, !is.na(fb_data))
total_fb_data <- nrow(df_facebook)
fb_data_pd_count <- count_individuals(df_facebook, pd_condition)
fb_data_et_count <- count_individuals(df_facebook, et_condition)
fb_data_parkinsonism_count <- count_individuals(df_facebook, parkinsonism_condition)
fb_data_caregivers_count <- count_individuals(df_facebook, caregiver_condition)

# People who never submitted their Facebook data
df_facebook_not_submitted <- subset(df_interviewed, is.na(fb_data) & has_facebook == 'Yes')
total_missing_fb_data <- nrow(df_facebook_not_submitted)

# People who participated without Facebook
df_nofacebook <- subset(df_interviewed, is.na(fb_data) & has_facebook == 'No')
no_fb_part_data <- nrow(df_nofacebook)
```

Generate overall enrollment flowchart
```{r}
# Install packages if not already installed
if (!require("DiagrammeR")) install.packages("DiagrammeR")
if (!require("DiagrammeRsvg")) install.packages("DiagrammeRsvg")
if (!require("rsvg")) install.packages("rsvg")

# Load the libraries
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)

# Create the flowchart
flowchart <- grViz(sprintf("
digraph G {
    node [shape = rectangle, style = filled, fillcolor = lightgrey]

    A [label = 'Total Attempts (n = %d)']
    B [label = 'Successful Contact (n = %d)']
    C [label = 'Agreed to Participate (n = %d)']
    D [label = 'Enrolled (n = %d)']
    E [label = 'Withdrawn (n = %d)']
    F [label = 'Completed Interview (n = %d)']
    G [label = 'Facebook User, Data Submitted (n = %d)']
    H [label = 'Facebook User, Data Not Submitted (n = %d)']
    I [label = 'Non-Facebook User (n = %d)']

    A -> B
    B -> C
    C -> D
    D -> E
    D -> F
    F -> G
    F -> H
    F -> I
}
", total_attempts, total_contact, total_agree, total_enroll, total_withdrew, total_interview, total_fb_data, total_missing_fb_data, no_fb_part_data))

flowchart
```

## Final sample by dx
```{r}
print("People interviewed")
print(paste("PD:", interview_pd_count))
print(paste("ET:", interview_et_count))
print(paste("AP:", interview_parkinsonism_count))
print(paste("Caregivers:", interview_caregivers_count))
```
```{r}
print("People interviewed who donated Facebook data")
print(paste("PD:", fb_data_pd_count))
print(paste("ET:", fb_data_et_count))
print(paste("AP:", fb_data_parkinsonism_count))
print(paste("Caregivers:", fb_data_caregivers_count))
```

## Percentage of contacted people with Facebook accounts
```{r}
# People who have FB, regardless of enrollment status
df_contacted_with_facebook <- df_success |> subset(has_facebook == 'Yes')
total_contacted_with_facebook <- nrow(df_contacted_with_facebook)
print(paste('Successfully contacted people with Facebooks:', total_contacted_with_facebook))
perc_facebook = total_contacted_with_facebook/total_contact
print(paste('Percentage of contacted people with FB:', round(100*perc_facebook,2), '%'))

# People without FB, regardless of enrollment status
df_contacted_no_facebook <- df_success |> subset(has_facebook == 'No')
total_contacted_no_facebook <- nrow(df_contacted_no_facebook)
print(paste('People without Facebooks:', total_contacted_no_facebook))
perc_facebook_no = total_contacted_no_facebook/total_contact
print(paste('Percentage of contacted people without FB:', round(100*perc_facebook_no,2), '%'))

# People with unknown FB, regardless of enrollment status
df_contacted_facebook_unknown <- df_success |> subset(is.na(has_facebook))
total_contacted_facebook_unknown <- nrow(df_contacted_facebook_unknown)
print(paste('People who didnt say whether they had Facebooks:', total_contacted_facebook_unknown))
perc_facebook_unknown = total_contacted_facebook_unknown/total_contact
print(paste('Percentage of contacted people who didnt say whether they had facebook:', round(100*perc_facebook_unknown,2), '%'))
```

## Percentage of contacted people with Facebook accounts who agreed to participate, regardless of whether they did
# Total number of people who agreed to participate
```{r}
df_contacted_with_facebook_agreed <- subset(df_contacted_with_facebook, contact_refused == 'No')
total_contacted_with_facebook_agreed <- nrow(df_contacted_with_facebook_agreed)
perc_contacted_with_facebook_agreed = total_contacted_with_facebook_agreed/total_contacted_with_facebook
print(paste('Percentage of contacted people who indicated they had a Facebook and agreed to participate:', round(100*perc_contacted_with_facebook_agreed,2), '%'))
```

## Demographic majorities
```{r}
# Compute counts and percentages
average_age = mean(df_interviewed$age)
sd_age = sd(df_interviewed$age)
print(paste('Average age of participants:', round(average_age[[1]], 2), '(SD = ', round(sd_age,2),')'))
```


```{r}
# Compute counts and percentages
gender_counts <- table(df_interviewed$gender)
gender_percent <- prop.table(gender_counts) * 100

# Print results
print(gender_counts)  # Absolute counts
print(round(gender_percent, 2))  # Percentages rounded to 2 decimal places
```

```{r}
# Compute counts and percentages
race_counts <- table(df_interviewed$race)
race_percent <- prop.table(race_counts) * 100

# Print results
print(round(race_percent, 2))  # Percentages rounded to 2 decimal places
```

```{r}
# Compute counts and percentages
ethnicity_counts <- table(df_interviewed$ethnicity)
ethnicity_percent <- prop.table(ethnicity_counts) * 100

# Print results
print(round(ethnicity_percent, 2))  # Percentages rounded to 2 decimal places
```

```{r}
# Compute counts and percentages
edu_counts <- table(df_interviewed$education)
edu_percent <- prop.table(edu_counts) * 100

# Print results
print(round(edu_percent, 2))  # Percentages rounded to 2 decimal places
print(paste('At least college educated: ', 40+38.33))
```

```{r}
library(arsenal)
# Generate the summary table
t1 <- tableby(dx_type ~ age + gender + race + ethnicity + education + disease_duration + mds_updrs_i_total + mds_updrs_ii_total + mds_updrs_iv_total, data = df_interviewed) |>
  summary(digits = 1, digits.pct = 0, digits.p = 2) |> 
  as.data.frame()

# Clean up the Variable column name
names(t1)[1] = "Variable"
t1$Variable = t1$Variable |> stringr::str_remove_all(stringr::fixed("**")) |> stringr::str_remove_all(stringr::fixed("&nbsp;"))

# Remove total row & p-value
#t1 = t1[,-6:-7]

# Remove PD variables for caregivers
#t1[18:36, 2] = ""

# Remove MDS UPDRS IV for ET patients
#t1[34:36, 3] = ""

t1 <- t1[!(t1$Variable == "N-miss" & t1$Variable %in% c("disease_duration")), ]

# Print the table
t1 |> knitr::kable() |> kableExtra::kable_styling()
write.csv(t1, "summary_table.csv", row.names = FALSE)
```



## Just people with PD
```{r}
df_pd = subset(df_interviewed, dx_type == "Parkinsons disease")

# Age
average_age = mean(df_pd$age)
sd_age = sd(df_pd$age)
print(paste('Average age of PwPD participants:', round(average_age[[1]], 2), '(SD = ', round(sd_age,2),')'))
```

```{r}
# Disease Duration
average_duration = mean(df_pd$disease_duration)
sd_duration = sd(df_pd$disease_duration)
print(paste('Average disease duration of PwPD participants:', round(average_duration[[1]], 2), '(SD = ', round(sd_duration,2),')'))
```

```{r}
# MDS UPDRS Part I
average <- mean(df_pd$mds_updrs_i_total, na.rm = TRUE)
sd = sd(df_pd$mds_updrs_i_total, na.rm = TRUE)
print(paste('Average MDS UPDRS Part I of PwPD participants:', round(average[[1]], 2), '(SD = ', round(sd,2),')'))

# MDS UPDRS Part II
average <- mean(df_pd$mds_updrs_ii_total, na.rm = TRUE)
sd = sd(df_pd$mds_updrs_ii_total, na.rm = TRUE)
print(paste('Average MDS UPDRS Part II of PwPD participants:', round(average[[1]], 2), '(SD = ', round(sd,2),')'))

# MDS UPDRS Part IV
average <- mean(df_pd$mds_updrs_iv_total, na.rm = TRUE)
sd = sd(df_pd$mds_updrs_iv_total, na.rm = TRUE)
print(paste('Average MDS UPDRS Part IV of PwPD participants:', round(average[[1]], 2), '(SD = ', round(sd,2),')'))
```


```{r}
## Just people with FB data
# Subset participants with Facebook data
df_facebook <- subset(df_interviewed, !is.na(fb_data))

# Check group sizes
table(df_facebook$dx_type)

# Compare age between PD and Caregiver groups
df_compare <- subset(df_facebook, dx_type %in% c("Parkinsons disease", "Caregivers"))

# Basic summary stats
library(dplyr)
df_compare %>%
  group_by(dx_type) %>%
  summarise(
    N = n(),
    Mean_Age = mean(age, na.rm = TRUE),
    SD_Age = sd(age, na.rm = TRUE)
  )

# T-test to compare age distributions
t.test(age ~ dx_type, data = df_compare)
```


